---
title: "merge_NaNDA_and_patient_data"
output: html_notebook
---

# Load libraries
```{r}
library(caret)
library(data.table)
library(dplyr)
library(sets)
library(scales)
library(tidyr)
library(stringr)
library(tibble)
library(tidygeocoder)
library(stringi)

library(readr) # For read_tsv
```

# Import raw NaNDA datasets: 
## NOTE: All at census tract level.
1.) National Neighborhood Data Archive (NaNDA): Broadband Internet Availability, Speed, and Adoption by Census Tract and ZIP Code Tabulation Area, United States, 2014-2020 (ICPSR 38567)
- 38567: 2014-2020

2.) National Neighborhood Data Archive (NaNDA): Education and Training Services by Census Tract, United States, 2003-2017
- 127681: 2003 – 2017 (2003-2015, 2017)

3.) National Neighborhood Data Archive (NaNDA): Internet Access by Census Tract and ZIP Code Tabulation Area, United States, 2015-2019 (ICPSR 38559)
- 38559: 2015-2019

4.) National Neighborhood Data Archive (NaNDA): Neighborhood-School Gap by Census Tract and ZIP Code Tabulation Area, United States, 2009-2010 and 2015-2016 (ICPSR 38579)
- 38579: 2009-2010, 2015-2016

5.) National Neighborhood Data Archive (NaNDA): Polluting Sites by Census Tract and ZIP Code Tabulation Area, United States, 1987-2021 (ICPSR 38597)
- 38597: 1987-2021

6.) National Neighborhood Data Archive (NaNDA): School District Characteristics and School Counts by Census Tract, ZIP Code Tabulation Area, and School District, 2000-2018 (ICPSR 38569)
- 38569: 2000-2018

7.) National Neighborhood Data Archive (NaNDA): Post Offices and Banks by Census Tract, United States, 2003-2017
- 128281: 2003-2018
```{r}

# 1.) 
broadband_df <- read_tsv("../NaNDA_total/NaNDA_Broadband_Internet/DS0001/38567-0001-Data.tsv")

# 2.) 
edtrain_df <- read.csv("../NaNDA_total/NaNDA_Education/nanda_edtrain_tract_2003-2017_01P.csv")

# 3.) 
internet_df <- read_tsv("../NaNDA_total/NaNDA_Internet_Access/DS0001/38559-0001-Data.tsv")

# 4.) 
schoolneigh_df <- read_tsv("../NaNDA_total/NaNDA_Neighborhood_School_Gap/DS0001/38579-0001-Data.tsv")

# 5.) V1.
polluting_df <- read_tsv("../NaNDA_total/NaNDA_Polluting_Sites/DS0001/38597-0001-Data.tsv")
  
# 6.) 
schooldist_df <-read_tsv("../NaNDA_total/NaNDA_School_District_Characteristics_and_School_Counts/DS0001/38569-0001-Data.tsv")


# 7.) CSV not TSV here.
postbanks_df <- read.csv("../NaNDA_total/NaNDA_Post_Offices_and_Banks/nanda_pobank_tract_2003-2017_02P_csv_with_readme/nanda_pobank_tract_2003-2017_02P.csv")



```

# Import NaNDA data:
- use 2 dataset that overlap with 2010-2018.
```{r}
# NaNDA: 16th dataset

# DS0001: 2010 census
# DS0003: 2020 census
#pollution_d16_2010 <- read_tsv("../NaNDA_total/NaNDA_Polluting_Sites/DS0001/38597-0001-Data.tsv")

# pollution_d16_2020 <- read_tsv("../NaNDA_total/NaNDA_Polluting_Sites/DS0003/38597-0003-Data.tsv")

#print(colnames(pollution_d16_2020))

# NaNDA: 22nd dataset

# school_d22 <- read_tsv("../NaNDA_total/NaNDA_School_District_Characteristics_and_School_Counts/DS0001/38569-0001-Data.tsv")

#print(colnames(school_d22))
```
# Pollution data: NaNDA
Fields: 
- TRACT_FIPS20 (11 digits): <chr>
- YEAR: <dbl>
- COUNT_TRI_FACILITIES: <dbl>
```{r}

# 2010-2018: with 'pollution_d16_2010'
# min(pollution_d16_2010$YEAR)
# max(pollution_d16_2010$YEAR)

# NOTE: We use the 2020 census data which ranges from 1987 to 2021 because it is the most up to date. 
# 1987 - 2021: with 'pollution_d16_2020'

#min(pollution_d16_2020$YEAR)
#max(pollution_d16_2020$YEAR)

#min(school_d22$YEAR)
# max(school_d22$YEAR)

```

# Read in patient data: tractmlm
```{r}

source('../../SU23 Rotation/sdoh_summer/aequitus/data/data_cleaners.R')

patient_df <- read.csv("../../SU23 Rotation/sdoh_summer/aequitus/data/tractmlm.csv")

# Before restricting to only black, white patients
nrow(patient_df)

# Restrict dataset to black, white patients only:
patient_df <- patient_df[(patient_df$black == 1 | patient_df$white == 1),]

nrow(patient_df)

############################
# Exclusion criteria; 
endpoints <- c("readmit30bin", "death30bin", "composite30")

# clean up the end-point related stuff
for (colx in endpoints) {
	patient_df <- remove_na(patient_df, colx)
}

## fix it to be adults aged >= 18
patient_df <- patient_df[(patient_df$age_current >= 18), ]

cat("After only adults: ", nrow(patient_df))
```


# Import AHRQ data:

# Load AHRQ data: county, census tract
```{r}

# County AHRQ
cty_ahrq_df_2010 <- read.csv("./COUNTY_AHRQ/SDOH_2010_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2011 <- read.csv("./COUNTY_AHRQ/SDOH_2011_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2012 <- read.csv("./COUNTY_AHRQ/SDOH_2012_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2013 <- read.csv("./COUNTY_AHRQ/SDOH_2013_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2014 <- read.csv("./COUNTY_AHRQ/SDOH_2014_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2015 <- read.csv("./COUNTY_AHRQ/SDOH_2015_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2016 <- read.csv("./COUNTY_AHRQ/SDOH_2016_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2017 <- read.csv("./COUNTY_AHRQ/SDOH_2017_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2018 <- read.csv("./COUNTY_AHRQ/SDOH_2018_COUNTY_1_1.csv") # header=FALSE

# Use fill to ensured consistent number of columns, merge dfs from 
# Merge dfs on the 6 items below in list.
cty_ahrq_total <- rbindlist(list(cty_ahrq_df_2010, cty_ahrq_df_2011, cty_ahrq_df_2012, cty_ahrq_df_2013,
                             cty_ahrq_df_2014, cty_ahrq_df_2015, cty_ahrq_df_2016, cty_ahrq_df_2017, cty_ahrq_df_2018), fill=TRUE)


# Save names WITHOUT suffix first.
cty_features_nosuffix <- colnames(cty_ahrq_total)
# Rename cty_ahrq_total FEATURE names: append "_countylevel"

# Columns to exclude from renaming
exclude_columns_cty <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "YEAR")

cty_ahrq_total <- cty_ahrq_total %>%
  rename_at(vars(-one_of(exclude_columns_cty)), ~paste0(., "_countylevel"))

# Import AHRQ data
tr_ahrq_df_2010 <- read.csv("./TRACT_AHRQ/sdoh_2010_tract_1_0.csv") # header=FALSE
tr_ahrq_df_2011 <- read.csv("./TRACT_AHRQ/sdoh_2011_tract_1_0.csv")
tr_ahrq_df_2012 <- read.csv("./TRACT_AHRQ/sdoh_2012_tract_1_0.csv")
tr_ahrq_df_2013 <- read.csv("./TRACT_AHRQ/sdoh_2013_tract_1_0.csv")
tr_ahrq_df_2014 <- read.csv("./TRACT_AHRQ/sdoh_2014_tract_1_0.csv")
tr_ahrq_df_2015 <- read.csv("./TRACT_AHRQ/sdoh_2015_tract_1_0.csv")
tr_ahrq_df_2016 <- read.csv("./TRACT_AHRQ/sdoh_2016_tract_1_0.csv")
tr_ahrq_df_2017 <- read.csv("./TRACT_AHRQ/sdoh_2017_tract_1_0.csv")
tr_ahrq_df_2018 <- read.csv("./TRACT_AHRQ/sdoh_2018_tract_1_0.csv")

tr_ahrq_total <- rbindlist(list(tr_ahrq_df_2010, tr_ahrq_df_2011, tr_ahrq_df_2012, tr_ahrq_df_2013,
                             tr_ahrq_df_2014, tr_ahrq_df_2015, tr_ahrq_df_2016, tr_ahrq_df_2017, tr_ahrq_df_2018), fill=TRUE)


# Save names WITHOUT suffix first.
trct_features_nosuffix <- colnames(tr_ahrq_total)


# Rename tr_ahrq_total FEATURE names: append "_census_tractlevel"
# Columns to exclude from renaming
exclude_columns_trct <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "TRACTFIPS", "TRACTFIPS_padded", "YEAR")

tr_ahrq_total <- tr_ahrq_total %>%
  rename_at(vars(-one_of(exclude_columns_trct)), ~paste0(., "_census_tractlevel"))



# SAVE VAR NAMES: 
# COUNTY:
cty_features <- colnames(cty_ahrq_total)
trct_features <- colnames(tr_ahrq_total)


all_gran_feats_metadata <- readxl::read_excel("2010-18-all-granularities-AHRQ-dict.xlsx") # has 2010-2018 feats, metadata.

# -- todo: merge all AHRQ features w/ their metadata (from cty_features_metadata ) 
# length(set_intersection(cty_features, cty_feats_metadata$`Variable Name`))

########### COUNTY LEVEL: #################
# COUNTY: 1267 county features (2010-2018)
length(set_intersection(cty_features_nosuffix, all_gran_feats_metadata$`Variable Name`))
# Only select feats metadata which are present in "cty_features."
cty_features_and_metadata <- all_gran_feats_metadata[all_gran_feats_metadata$`Variable Name` %in% cty_features_nosuffix, ]

# Lower-case domain names.
cty_features_and_metadata <- cty_features_and_metadata %>%
  mutate(Domain = stri_trans_tolower(stri_trans_general(Domain, "Latin-ASCII")))

# Columns to exclude from renaming
exclude_columns_cty <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "YEAR")

cty_features_and_metadata <- cty_features_and_metadata %>%
  mutate(`Variable Name` = case_when(
    `Variable Name` %in% exclude_columns_cty ~ `Variable Name`,
    TRUE ~ paste0(`Variable Name`, "_countylevel")
  ))


# Print the modified dataframe
print(cty_features_and_metadata)


########### TRACT LEVEL: #################
length(set_intersection(trct_features_nosuffix, all_gran_feats_metadata$`Variable Name`))
# TRACT: 519 county features (2010-2018)
# Only select feats metadata which are present in "trct_features_nosuffix"
trct_features_and_metadata <- all_gran_feats_metadata[all_gran_feats_metadata$`Variable Name` %in% trct_features_nosuffix, ]

# Lower-case domain names.
trct_features_and_metadata <- trct_features_and_metadata %>%
  mutate(Domain = stri_trans_tolower(stri_trans_general(Domain, "Latin-ASCII")))

# Append _census_tractlevel to all feats in this METADATA:

exclude_columns_trct <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "TRACTFIPS", "YEAR") # "TRACTFIPS_padded"

trct_features_and_metadata <- trct_features_and_metadata %>%
  mutate(`Variable Name` = case_when(
    `Variable Name` %in% exclude_columns_trct ~ `Variable Name`,
    TRUE ~ paste0(`Variable Name`, "_census_tractlevel")
  ))


# Print the modified dataframe
print(trct_features_and_metadata)
```

# Save AHRQ census tract variables for LLM annotation:
```{r}
llm_tract_AHRQ_prompts <- trct_features_and_metadata

# 1.) Rename columns: 
# Domain, Variable Name, Variable Label, Data Source
# domain, variable_name, variable_label, data_source
llm_tract_AHRQ_prompts <- llm_tract_AHRQ_prompts %>%
  rename(
    domain = Domain,
    SDOH_topic = Topic,
    variable_name = `Variable Name`,  # Enclose the column name in backticks because it contains spaces
    variable_label = `Variable Label`, # Enclose the column name in backticks because it contains spaces
    data_source = `Data Source`       # Enclose the column name in backticks because it contains spaces
  )


# 2.) Don't include vars: drop these rows.
exclude_columns_trct <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "TRACTFIPS", "YEAR") # "TRACTFIPS_padded"
llm_tract_AHRQ_prompts <- llm_tract_AHRQ_prompts %>%
  filter(!(variable_name %in% exclude_columns_trct))

# We *could* insert full names in data_source:

write.csv(llm_tract_AHRQ_prompts ,"AHRQ_tract_varnames_2010-2018.csv", row.names = FALSE )

nrow(llm_tract_AHRQ_prompts)


```


```{r}
# SVI Year: 2014. SVI Geo level: census tract. All states in US included. 
SVI_2014_US <- read.csv("./TRACT_AHRQ/SVI_2014_US.csv") 

keep_svi <- c(
              "FIPS",
              "SPL_THEMES",
              "EPL_POV", "EPL_UNEMP", "EPL_PCI", "EPL_NOHSDP",
              "EPL_AGE65", "EPL_AGE17", "EPL_DISABL", "EPL_SNGPNT",
              "EPL_MINRTY", "EPL_LIMENG", "EPL_MUNIT", "EPL_MOBILE",
              "EPL_CROWD", "EPL_NOVEH", "EPL_GROUPQ",
              "RPL_THEME1",
              "RPL_THEME2",
              "RPL_THEME3",
              "RPL_THEME4",
              "EPL_POV", "EPL_UNEMP", "EPL_PCI", "EPL_NOHSDP",
              "EPL_AGE65", "EPL_AGE17", "EPL_DISABL", "EPL_SNGPNT",
              "EPL_MINRTY", "EPL_LIMENG", "EPL_MUNIT", "EPL_MOBILE",
              "EPL_CROWD", "EPL_NOVEH", "EPL_GROUPQ",
              "RPL_THEMES",
              "F_TOTAL",
              "E_HU", "E_HH", "E_POV", "E_UNEMP", "E_PCI", 
                   "E_NOHSDP", "E_AGE65", "E_AGE17", "E_DISABL", "E_SNGPNT", 
                   "E_MINRTY", "E_LIMENG", "E_MUNIT", "E_MOBILE", "E_CROWD", 
                   "E_NOVEH", "E_GROUPQ",
              "EP_POV", "EP_UNEMP", "EP_PCI", "EP_NOHSDP", 
                            "EP_AGE65", "EP_AGE17", "EP_DISABL", "EP_SNGPNT", 
                            "EP_MINRTY", "EP_LIMENG", "EP_MUNIT", "EP_MOBILE", 
                            "EP_CROWD", "EP_NOVEH", "EP_GROUPQ")

# Subset only necessary SVI columns + FIPS (tract fips). Don't need county fips.
SVI_2014_US <- SVI_2014_US[, keep_svi]

head(SVI_2014_US)

```


# Drop rows missing from patient data: 
```{r}
patient_df <- patient_df %>% drop_na(censustract)
patient_df <- patient_df %>% drop_na(tractnum)

patient_df <- patient_df %>% drop_na(countynum)
patient_df <- patient_df %>% drop_na(statenum)

###
nrow(patient_df)
```


# Cast Census Tract (i.e., censustract variable), 
  & County Fips (i.e., fips variable) to proper length for merge with AHRQ (keep these 2 cols as type *chr*)
  
---- NOTE: 
- Tract fips: always should be 11 digits.
- County fips: always should be 5 digits.

```{r}

########################
# Patient data: cast FIPS values
# County: to 5 chars. Note: 'fips' has state[2 chr]+county[3 chr]
#patient_df$fips
patient_df$fips <- sprintf("%05s", patient_df$fips)

# Tract: to 11 chars
#patient_df$censustract
patient_df$censustract <- sprintf("%011s", patient_df$censustract)

########################
# 1.) AHRQ: cast FIPS values
cty_ahrq_total$COUNTYFIPS_padded <- sprintf("%05s", cty_ahrq_total$COUNTYFIPS)

tr_ahrq_total$TRACTFIPS_padded <- sprintf("%011s", tr_ahrq_total$TRACTFIPS)

########################
# 2.) SVI_2014_US
# Cast FIPS to 11 digit <chr> (before, some are 10, some are 11).
SVI_2014_US$FIPS <- sprintf("%011s", SVI_2014_US$FIPS)


########################
# 3.) NaNDA datasets: 


# Cast: census tract FIPS for all.
#pollution_d16_2020$TRACT_FIPS20 <- sprintf("%011s", pollution_d16_2020$TRACT_FIPS20)
#school_d22$TRACT_FIPS10 <- sprintf("%011s", school_d22$TRACT_FIPS10)

# 7 Datasets for MLHC:
# 1.) 
broadband_df$TRACT_FIPS10 <- sprintf("%011s", broadband_df$TRACT_FIPS10)
# 2.) CSV not TSV
edtrain_df$tract_fips10 <- sprintf("%011s", edtrain_df$tract_fips10)
# 3.) 
internet_df$TRACT_FIPS10 <- sprintf("%011s", internet_df$TRACT_FIPS10)
# 4.) 
schoolneigh_df$TRACT_FIPS10 <- sprintf("%011s", schoolneigh_df$TRACT_FIPS10)
# 5.) V4.
polluting_df$TRACT_FIPS10 <-  sprintf("%011s", polluting_df$TRACT_FIPS10)
#  6.) 
schooldist_df$TRACT_FIPS10 <- sprintf("%011s", schooldist_df$TRACT_FIPS10)
# 7.) CSV not TSV here.
postbanks_df$tract_fips10 <- sprintf("%011s", postbanks_df$tract_fips10)


# NaNDA: rename [geo & year] columns --> CAPITAL letters:

# 2.) CSV not TSV
colnames(edtrain_df)[colnames(edtrain_df) == "tract_fips10"] <- toupper("tract_fips10")
colnames(edtrain_df)[colnames(edtrain_df) == "year"] <- toupper("year")

# 7.) CSV not TSV here.
colnames(postbanks_df)[colnames(postbanks_df) == "tract_fips10"] <- toupper("tract_fips10")
colnames(postbanks_df)[colnames(postbanks_df) == "year"] <- toupper("year")


# ALL NaNDA: Rename all FEATURE columns to be upper case.
colnames(broadband_df) <- toupper(colnames(broadband_df))
colnames(edtrain_df) <- toupper(colnames(edtrain_df))
colnames(internet_df) <- toupper(colnames(internet_df))
colnames(schoolneigh_df) <- toupper(colnames(schoolneigh_df))
colnames(polluting_df) <- toupper(colnames(polluting_df))
colnames(schooldist_df) <- toupper(colnames(schooldist_df))
colnames(postbanks_df) <- toupper(colnames(postbanks_df))

```

# NaNDA: Impute: 
- approach 1: For each variable V, impute most recent value (past year Y-1) of variable V in census tract C.
- approach 2: If no past years, impute the value using the closest year of V in census tract C.
```{r}
# Define a function to impute missing values based on the closest non-missing value from a previous year
impute_missing_helper <- function(x) {
  x_filled <- zoo::na.locf(x, fromLast = TRUE, na.rm = FALSE)
  x_filled <- zoo::na.locf(x_filled, na.rm = FALSE)
  # Ensure the imputed variable has the same size as the original variable
  # Find the NA values that were not imputed
  return(x_filled)
}

# Assumes we have TRACT_FIPS10, YEAR
impute_df <- function(imp_df) {
  # Sort imp_df
  imp_df <- imp_df %>%
  arrange(TRACT_FIPS10, YEAR)
  # Impute imp_df with helper
  imp_df <- imp_df %>%
  group_by(TRACT_FIPS10) %>%
  mutate_at(vars(-YEAR, -TRACT_FIPS10), impute_missing_helper)
  return(imp_df)
}


# Impute missing values for each dataset:
broadband_df <- impute_df(broadband_df)
edtrain_df <- impute_df(edtrain_df)
# internet_df <- impute_df(internet_df) # Only has 1 year, not many. The ACS-derived data contains one measure for the entire period of 2015-2019.
schoolneigh_df <- impute_df(schoolneigh_df)
polluting_df <- impute_df(polluting_df)
schooldist_df <- impute_df(schooldist_df)
postbanks_df <- impute_df(postbanks_df)
```

# Check NaNDA imputation results:
```{r}

sum(is.na(broadband_df$RFHSC_PER_1000_HHS)) # Only 38 values still missing across all tracts & years

sum(is.na(edtrain_df$aland10)) # 14028 values still missing across all tracts & years. These are from Puerto Rico, Virgin Islands etc...
missing_rates <- broadband_df %>%
  group_by(TRACT_FIPS10) %>%
  summarize(
    missing_rate_A = mean(is.na(RFHSC_PER_1000_HHS_10_1))
  )
sum(missing_rates$missing_rate_A == 1)
# --> Tells us there are 19 census tracts that have NO values over 2014 - 2020 for the variables RFHSC_PER_1000_HHS, RFHSC_PER_1000_HHS_10_1.
# If patients live in these census tracts, we will drop these patients.

```

# AHRQ: Impute
# ###########################################
```{r}
library(rlang)

# Define a function to impute missing values based on the closest non-missing value from a previous year
impute_missing_helper <- function(x) {
  x_filled <- zoo::na.locf(x, fromLast = TRUE, na.rm = FALSE)
  x_filled <- zoo::na.locf(x_filled, na.rm = FALSE)
  # Ensure the imputed variable has the same size as the original variable
  # Find the NA values that were not imputed
  return(x_filled)
}

# Assumes we have TRACT_FIPS10, YEAR
impute_df <- function(imp_df, yr_var, geo_var) {
  # Sort imp_df
  imp_df <- imp_df %>%
  arrange(!!sym(geo_var), !!sym(yr_var))
  # Impute imp_df with helper
  imp_df <- imp_df %>%
  group_by(!!sym(geo_var)) %>%
  mutate_at(vars(-!!sym(yr_var), -!!sym(geo_var)), impute_missing_helper)
  return(imp_df)
}

# Impute county.
cty_ahrq_total <- impute_df(cty_ahrq_total, 'YEAR', 'COUNTYFIPS_padded') # "COUNTYFIPS_padded", "YEAR"
# Impute tract.
tr_ahrq_total <- impute_df(tr_ahrq_total, 'YEAR', 'TRACTFIPS_padded') # "TRACTFIPS_padded", "YEAR"

```

# Check NaNDA imputation results:
```{r}
sum(is.na(cty_ahrq_total$ACS_TOT_POP_WT_countylevel)) # Only 38 values still missing across all tracts & years


```


# Merge: 
- tractmlm: patient_df
SDOH data sources: 
1+ ahrq: 
  - cty_ahrq_total
  - tr_ahrq_total 
  
2+ SVI_2014_US: 
3+ NaNDA: (7) datasets. [capital or lowercase]
```{r}
# MERGE: both on subject's 'baselineyear' (admission year)
# 1.) Bring in County AHRQ, Census Tract AHRQ:
nrow(cty_ahrq_total)
sdoh_baseline <- merge(patient_df, cty_ahrq_total, by.x=c("fips", "baselineyear"), by.y=c("COUNTYFIPS_padded", "YEAR"), all.x = TRUE)
# nrow(sdoh_baseline_cty)
```

```{r}
# Tract AHRQ baseline: use later for patient-level cenus tract missingness calculations. 
sdoh_baseline <- merge(sdoh_baseline, tr_ahrq_total, by.x=c("censustract", "baselineyear"), by.y=c("TRACTFIPS_padded", "YEAR"), all.x = TRUE)
# nrow(sdoh_baseline_cty)
```


```{r}
# Next, merge in SVI 2014 data:
# use cty code: sdoh_baseline ++ SVI_2014_US (by 'censustract', 'FIPS')

# 2.) Bring in 2014 SVI by Census Tract:
sdoh_baseline <- merge(sdoh_baseline, SVI_2014_US  , by.x="censustract", by.y="FIPS", all.x = TRUE )
nrow(sdoh_baseline)
```



```{r}
# 3.) Bring in (7) NaNDA datasets: 1 at a time.
# census tract level.

# 1.) 
sdoh_baseline <- merge(sdoh_baseline, broadband_df  , by.x=c("censustract", "baselineyear"), by.y=c("TRACT_FIPS10", "YEAR"), all.x = TRUE )
# 2.) 
sdoh_baseline <- merge(sdoh_baseline, edtrain_df  , by.x=c("censustract", "baselineyear"), by.y=c("TRACT_FIPS10", "YEAR"), all.x = TRUE )
# 3.) 
# Only has 1 year, not many. The ACS-derived data contains one measure for the entire period of 2015-2019.
sdoh_baseline <- merge(sdoh_baseline, internet_df  , by.x=c("censustract"), by.y=c("TRACT_FIPS10"), all.x = TRUE )
# 4.) 
sdoh_baseline <- merge(sdoh_baseline, schoolneigh_df  , by.x=c("censustract", "baselineyear"), by.y=c("TRACT_FIPS10", "YEAR"), all.x = TRUE )
# 5.) 
sdoh_baseline <- merge(sdoh_baseline, polluting_df  , by.x=c("censustract", "baselineyear"), by.y=c("TRACT_FIPS10", "YEAR"), all.x = TRUE )
#  6.) 
sdoh_baseline <- merge(sdoh_baseline, schooldist_df  , by.x=c("censustract", "baselineyear"), by.y=c("TRACT_FIPS10", "YEAR"), all.x = TRUE )
# 7.) CSV not TSV here.
sdoh_baseline <- merge(sdoh_baseline, postbanks_df  , by.x=c("censustract", "baselineyear"), by.y=c("TRACT_FIPS10", "YEAR"), all.x = TRUE )


nrow(sdoh_baseline)
```


```{r}
head(sdoh_baseline) # Patient data + matched tract.
nrow(sdoh_baseline)
```


# Before feeding to classifiers, drop potential PHI from sdoh_baseline
```{r}


remove_cols <- c("patient_name", "Patient",  "Patient_Address",
                 "", "Patient_Home_Street_Address", "Patient_Home_City", "Patient_Home_State", "dob",
                 "Patient_Death_Day")

drop_vars <- c("patient_name", "Patient_Home_Street_Address", "Patient", 
               "Patient_Address", "Match_Address", "Patient_Birth_Date", "Patient_Home_City", "Patient_Home_State",
               "dob", "Patient_Death_Day",
               "Lon__Lat", "cxy_address", "cxy_matched_address", "cxy_tiger_line_id", "cxy_tract_id", "cxy_block_id")
# Remove drop_vars
sdoh_baseline <- sdoh_baseline[, !(names(sdoh_baseline) %in% drop_vars)]

```




# Build subgroup variables.
```{r}
sdoh_baseline <- sdoh_baseline %>%
  mutate(baselineyear = as.numeric(baselineyear), na.rm=FALSE)

############ Prepare subgroup flags:
# Subgroups to eval on: 
# 1. Race
# white <int>: 0/1
# black <int>: 0/1
# other <int>: 0/1
sdoh_baseline <- sdoh_baseline %>%
  mutate(blackorwhite = if_else(black == 1 | white == 1, 1, 0))

# 2. Sex (or binary gender)
# female <int>: 0/1
# ADD this field: male <int>: 0/1
sdoh_baseline <- sdoh_baseline %>%
  mutate(male = if_else(female == 0, 1, 0))

# 3. The 6 subgroup combinations: 
# c(white_male, black_male, other_male, white_female, black_female, other_female)

# Note that all 6 are mutually exclusive here.
sdoh_baseline <- sdoh_baseline %>%
  mutate(white_male = if_else(white == 1 & female == 0, 1, 0))
sdoh_baseline <- sdoh_baseline %>%
  mutate(black_male = if_else((black == 1 & female == 0), 1, 0))
sdoh_baseline <- sdoh_baseline %>%
  mutate(other_male = if_else(other == 1 & female == 0, 1, 0))
sdoh_baseline <- sdoh_baseline %>%
  mutate(white_female = if_else(white == 1 & female == 1, 1, 0))
sdoh_baseline <- sdoh_baseline %>%
  mutate(black_female = if_else(black == 1 & female == 1, 1, 0))
sdoh_baseline <- sdoh_baseline %>%
  mutate(other_female = if_else(other == 1 & female == 1, 1, 0))



#######################
# SUBGROUP FOR AGEs: 
quarts <- quantile(sdoh_baseline$age_current, probs = seq(0, 1, 0.25))
upper_Q1_age <- quarts[2]
upper_Q2_age <- quarts[3]
upper_Q3_age <- quarts[4]
upper_Q4_age <- quarts[5]

sdoh_baseline <- sdoh_baseline %>%
  mutate(age_Q1 = if_else(age_current <= upper_Q1_age, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(age_Q2 = if_else(age_current <= upper_Q2_age & age_current > upper_Q1_age, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(age_Q3 = if_else(age_current <= upper_Q3_age & age_current > upper_Q2_age, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(age_Q4 = if_else(age_current <= upper_Q4_age & age_current > upper_Q3_age, 1, 0))



#######################
# HF: Sex & age quartiles
#######################
# Hypertension
sdoh_baseline <- sdoh_baseline %>%
  mutate(no_htn = if_else(htn == 0, 1, 0))

#######################
# Coronary Artery Disease
sdoh_baseline <- sdoh_baseline %>%
  mutate(no_cad = if_else(cad == 0, 1, 0))

#######################
# Chronic Kidney Disease
sdoh_baseline <- sdoh_baseline %>%
  mutate(no_ckd = if_else(ckd == 0, 1, 0))


#######################
# Charlson Comorbidity Index
  # Theoretical Max CCI: 24 (updated version) or 29 (older version).
  # Our dataset: Max CCI = 22.
CCI_quarts <- quantile(sdoh_baseline$cci, probs = seq(0, 1, 0.25))

# 1.) CCI steps 0-6, 7+ groups.
# 


sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_0 = if_else(cci == 0, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_1 = if_else(cci == 1, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_2 = if_else(cci == 2, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_3 = if_else(cci == 3, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_4 = if_else(cci == 4, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_5 = if_else(cci == 5, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_6 = if_else(cci == 6, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_7up = if_else(cci >= 7, 1, 0))





# 2.) CCI quartiles (second way to evaluate CCI)
upper_Q1_CCI <- CCI_quarts[2]
upper_Q2_CCI <- CCI_quarts[3]
upper_Q3_CCI <- CCI_quarts[4]
upper_Q4_CCI <- CCI_quarts[5]


sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_Q1 = if_else(cci <= upper_Q1_CCI, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_Q2 = if_else(cci <= upper_Q2_CCI & cci > upper_Q1_CCI, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_Q3 = if_else(cci <= upper_Q3_CCI & cci > upper_Q2_CCI, 1, 0))

sdoh_baseline <- sdoh_baseline %>%
  mutate(CCI_Q4 = if_else(cci > upper_Q3_CCI, 1, 0))
  
# TODO: Make feat set of AHRQ total + NaNDA total.
# TODO: adjust variable names!! Make all NaNDA vars *lowercase.*
# 3.) education = upper case



nrow(sdoh_baseline)
```


```{r}
############

sdoh_baseline <- create_folds(as.data.frame(sdoh_baseline), endpoints) # NOTE: For nested k-fold CV.


write.csv(sdoh_baseline, file="./MLHC_modelinputs.csv", row.names=FALSE) 
#old: clean_tract_ahrq_AQIimp1_baseline

#table(sdoh_baseline$readmit30bin_folds)
#table(sdoh_baseline$death30bin_folds)
#table(sdoh_baseline$composite30_folds)

# Ex run: 80-20 train-test split --> then k-fold s
#set.seed(42)
#sample <- sample(c(TRUE, FALSE), nrow(sdoh_baseline), replace=TRUE, prob=c(0.8,0.2))
#train  <- sdoh_baseline[sample, ]
#test   <- sdoh_baseline[!sample, ]


# twoormore_gap
# 36% of patients have this NaNDA feat missing even after "closest year" imp... (ADEN_611410)
sum(is.na(sdoh_baseline$ADEN_611410))/nrow(sdoh_baseline) # COUNT_EMP_522110

```

# Check input file variables:
```{r}
mvd_input_total_feats <- read.csv(file="../../SU23 Rotation/sdoh_summer/private_repo/data/MLHC Total/MLHC_modelinputs.csv")


sum(is.na(mvd_input_total_feats$COUNT_SALES_6111))

sum(is.na(mvd_input_total_feats$POPULATION.x))

sum(is.na(mvd_input_total_feats$POPULATION.y))


```



# ###############################
# Table 1. SUMMARY STATS
# Read in model input data
```{r}
input_total_feats <- read.csv(file="./MLHC_modelinputs.csv")


# Check readmit30bin positive rates by "race" subgroup, fold.
input_total_feats %>% group_by(Race, readmit30bin_folds) %>% summarize(newvar = mean(readmit30bin, na.rm = TRUE))
########################
```



```{r}
## tableone package itself
library(tableone)
library(kableExtra)

# Make strata column: white, black, other (1,2,3)
# make using cols:
# Drop "Race" == "other"




input_total_feats <- input_total_feats <- transform(
  input_total_feats, "Race"= ifelse(white==1, "White", ifelse(black==1, "Black", NA)))
  

# Make insurance column: "privinsurance","Medicare", "Medicaid",
input_total_feats <- input_total_feats <- transform(
  input_total_feats, "Insurance_type"= ifelse(privinsurance==1, "Private", ifelse(Medicare==1, "Medicare", ifelse(Medicaid==1, "Medicaid", "Not recorded"))))
  

# Make HF type: hfsystolic, hfdiastolic, hfother
input_total_feats <- input_total_feats <- transform(
  input_total_feats, "HF_classification"= ifelse(hfsystolic==1, "Systolic (HFrEF)", ifelse(hfdiastolic==1, "Diastolic (HFpEF)", ifelse(hfother==1, "Other", "Not recorded"))))



# vars to summarize (i.e demographics)
# htn: Hypertension
# diabetes: ___not available___
# cad: Coronary Artery Disease
# ckd: Chronic Kidney Disease
# sdi: Social deprivation index score
# ADI_national_rank: 
# ADI_state_rank: 
# svi: i.e. RPL_THEMES
# cci: Charlson Comorbidity Index
# Systolic_BP: Systolic blood pressure, mm Hg
# Heart_Rate: Heart Rate, beats per minute
# Resp: Respiratory rate, breaths per minute

# Rename feats not needing spaces: 
colnames(input_total_feats)[colnames(input_total_feats) == "age_current"] ="Age, years"
colnames(input_total_feats)[colnames(input_total_feats) == "age_Q1"] ="Age, Q1"
colnames(input_total_feats)[colnames(input_total_feats) == "age_Q2"] ="Age, Q2"
colnames(input_total_feats)[colnames(input_total_feats) == "age_Q3"] ="Age, Q3"
colnames(input_total_feats)[colnames(input_total_feats) == "age_Q4"] ="Age, Q4"

colnames(input_total_feats)[colnames(input_total_feats) == "female"] = "Female"
colnames(input_total_feats)[colnames(input_total_feats) == "male"] = "Male"

colnames(input_total_feats)[colnames(input_total_feats) == "black"] = "Black"
colnames(input_total_feats)[colnames(input_total_feats) == "white"] = "White"

colnames(input_total_feats)[colnames(input_total_feats) == "htn"] = "Hypertension"
colnames(input_total_feats)[colnames(input_total_feats) == "cad"] = "Coronary Artery Disease"
colnames(input_total_feats)[colnames(input_total_feats) == "ckd"] = "Chronic Kidney Disease"
colnames(input_total_feats)[colnames(input_total_feats) == "cci"] = "Charlson Comorbidity Index"
colnames(input_total_feats)[colnames(input_total_feats) == "Systolic_BP"] = "Systolic blood pressure, mm Hg"
colnames(input_total_feats)[colnames(input_total_feats) == "Heart_Rate"] = "Heart Rate, beats per minute"
colnames(input_total_feats)[colnames(input_total_feats) == "Resp"] = "Respiratory rate, breaths per minute"
colnames(input_total_feats)[colnames(input_total_feats) == "sdi"] = "Social Deprivation Index (percentile)"
colnames(input_total_feats)[colnames(input_total_feats) == "ADI_national_rank"] = "Area Deprivation Index, national rank (percentile)"
colnames(input_total_feats)[colnames(input_total_feats) == "ADI_state_rank"] = "Area Deprivation Index, state rank (percentile)"
colnames(input_total_feats)[colnames(input_total_feats) == "RPL_THEMES"] = "Social Vulnerability Index score (composite score)"

# New feats:
#colnames(input_total_feats)[colnames(input_total_feats) == "readmit30bin"] = "30-day Binary HF Readmission"
#colnames(input_total_feats)[colnames(input_total_feats) == "death30bin"] = "30-day Binary HF Death"

# Referring specialties: 
colnames(input_total_feats)[colnames(input_total_feats) == "specialcvd"] = "CVD Specialist" 
colnames(input_total_feats)[colnames(input_total_feats) == "specialint"] = "Int. Med. Specialist"
colnames(input_total_feats)[colnames(input_total_feats) == "specialeme"] = "EME Specialist"
colnames(input_total_feats)[colnames(input_total_feats) == "specialoth"] = "Other Specialist"

# Location treated: 
colnames(input_total_feats)[colnames(input_total_feats) == "loceuhfloor"] = "UH Floor" 
colnames(input_total_feats)[colnames(input_total_feats) == "loceclhicu"] = "CLH ICU"
colnames(input_total_feats)[colnames(input_total_feats) == "loceclhfloor"] = "CLH Floor"
colnames(input_total_feats)[colnames(input_total_feats) == "loceuhicu"] = "UH ICU"

# SDOH feats (comp -- other Circulation HF paper)

colnames(input_total_feats)[colnames(input_total_feats) == "percnt_ltfpl100"] = "percnt_ltfpl100" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_singlparntfly"] = "percnt_singlparntfly" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_dropout"] = "percnt_dropout" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_rentoccup"] = "percnt_rentoccup" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_crowding"] = "percnt_crowding" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_nonemp"] = "percnt_nonemp" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_unemp"] = "percnt_unemp" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_highneeds"] = "percnt_highneeds" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_hispanic"] = "percnt_hispanic" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_frgnborn"] = "percnt_frgnborn" 
colnames(input_total_feats)[colnames(input_total_feats) == "percnt_black"] = "percnt_black" 
colnames(input_total_feats)[colnames(input_total_feats) == "LOScheck"] = "Length of stay" 

# sdi
# ADI_state_rank
# ADI_national_rank


# Rename feats with spaces: 
colnames(input_total_feats)[colnames(input_total_feats) == "Insurance_type"] = "Insurance type"
colnames(input_total_feats)[colnames(input_total_feats) == "HF_classification"] = "HF classification"




# Length of stay, days: LOScheck
#myVars <- c("Age, years", "female", "Insurance type", "HF classification", "htn", "cad", "ckd", "cci", 
#            "Systolic_BP", "Heart_Rate", "Resp", 
#            "LOScheck", "sdi", "ADI_national_rank", "ADI_state_rank", "RPL_THEMES")

myVars <- c("Age, years","Age, Q1","Age, Q2","Age, Q3","Age, Q4", "Male", "Female", "Black", "White",
            "CVD Specialist", "Int. Med. Specialist", "EME Specialist", "Other Specialist",
            "UH Floor", "CLH ICU", "CLH Floor", "UH ICU",
            
            "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease", 
            "Charlson Comorbidity Index", 
            "Systolic blood pressure, mm Hg", "Heart Rate, beats per minute", "Respiratory rate, breaths per minute", 
            "Length of stay", "Social Deprivation Index (percentile)", "Area Deprivation Index, national rank (percentile)", "Area Deprivation Index, state rank (percentile)", "Social Vulnerability Index score (composite score)",
            "percnt_ltfpl100", "percnt_singlparntfly", "percnt_dropout", "percnt_rentoccup", "percnt_crowding", "percnt_nonemp", "percnt_unemp", "percnt_highneeds", "percnt_hispanic", "percnt_frgnborn", "percnt_black"
            
            )


## Vector of categorical variables that need transformation
catVars <- c("Women", "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease")

tab1 <- CreateTableOne(vars = myVars, strata = "readmit30bin", data = input_total_feats, addOverall = TRUE) 
# For race strata instead of outcome strata: strata = "Race"
# For p-values: test=TRUE, smd = TRUE

# typeof(tab1["test"])



# If you just say nonnormal = TRUE, all variables are summarized the “nonnormal” way.
#print(tab1, formatOptions = list(big.mark = ","))
# kable(p, format = "latex")
p <- print(tab1, printToggle = FALSE, noSpaces = TRUE)
p <- p[,1:4] # Remove "test" column.

kable(p, format = "latex", booktabs = TRUE)
```




